\section{Problem Context}

\subsection{Scenario and Pain Points}

To motivate the problem, we begin with a concrete collaborative scenario commonly observed in university projects.\par
University students frequently work in temporary groups and need to exchange files efficiently without relying on complex infrastructure, cloud account management, or external storage devices. During the course of a project, team members collaboratively produce and modify heterogeneous files, including source code, documents, images, datasets, and presentation materials. Since collaborators often use different operating systems and development environments, cross-platform compatibility further increases the complexity of file sharing and coordination.\par
In practice, temporary student teams usually lack the time, resources, and technical expertise required to maintain a dedicated file management or synchronization system. As a result, files are often exchanged through ad hoc communication channels such as group chats or email attachments. This approach introduces several issues, including duplicated files, inconsistent versions, lack of synchronization, and absence of proper change tracking.\par
In addition, ownership and persistence of project data become problematic once the collaboration ends. Some members may wish to preserve the project files for future reference, while others may prefer to discard them. However, because the files are typically distributed informally and managed inconsistently, long-term access often depends on a single individual maintaining the archive. This creates risks of data loss, incomplete project history, and fragmented ownership.\par
Therefore, the students require a lightweight and decentralized solution that enables them to:
\begin{itemize}
    \item create a shared repository without complicated server deployment or administrative setup;
    \item support synchronization of heterogeneous file types across multiple platforms;
    \item track modifications and maintain version history;
    \item provide shared ownership of project data among collaborators; and
    \item avoid dependence on a single participant for long-term storage, archival, or deletion of the project files.
\end{itemize}

\subsection{Problem Definition}

Based on the scenario described above, small-scale collaborative groups require a file-sharing mechanism that is lightweight, decentralized, and easy to use. In such environments, participants mainly care about whether they can quickly form a shared workspace, exchange project files reliably, preserve previous versions when changes are made, and continue working with the shared folder as if it were a normal local directory.\par
From this objective, the problem can be defined as the design of a peer-to-peer directory-level synchronization system with lightweight version control. More specifically, the system must solve the following technical problems:

% TODO: After finishing the later chapter, I feel that we need to adjust the following structure
% The main dimention should be:
% - Membership Management
% - Synchronization
% - Version Control 
% - Transparent Filesystem Integration (UX important)
% - File support
% And we will have sub-goal for part of it:
% - Rapid Propagation
% - Large File Support
% - Dynamic Membership
\begin{itemize}

\item \textbf{Membership Management.}

The system must allow collaborators to discover peers,
establish sharing relationships, and manage participation in
a shared workspace without relying on centralized account
management. It should support dynamic membership, allowing
peers to join, leave, or temporarily disconnect during
collaboration.

\item \textbf{Peer-to-Peer Synchronization}

The system must automatically propagate updates among
participating peers without requiring manual file exchange.
It should support multi-peer collaboration, tolerate
temporary peer disconnections, and maintain a reasonably
consistent shared workspace during active collaboration.

\item \textbf{Version Management.}

The system must provide lightweight version management and
rollback capabilities. Before a synchronized update
overwrites an existing file, an earlier state should be
preserved to allow recovery from accidental modifications,
incorrect updates, or synchronization errors.

\item \textbf{Heterogeneous File Support.}

The system must support heterogeneous project artifacts,
including source code, documents, images, multimedia content,
datasets, and other binary files. In addition, large files
should be transferred efficiently without assuming that all
data can be exchanged through small messages.

\item \textbf{Transparent Filesystem Integration.}

The shared workspace should behave like a normal directory on
the user's computer. Users should be able to create, edit,
move, and delete files using their existing tools, while
synchronization and version management are handled
transparently in the background.

\end{itemize}

Therefore, the central technical problem is to design a
serverless peer-to-peer directory synchronization system that
provides effective membership management, automatic
synchronization, lightweight history management, support for
heterogeneous project files, and transparent filesystem
integration while preserving the simplicity of ordinary local
folder-based collaboration.
